Run-Time Monitoring with Adjustable Overheads Using Dataflow-Guided Filtering

### 2014-09-12 evaluation.tex
Updated evaluation and figures.
Changed figures to use size 8 Arial.
Added LockSet data.
Added min/max whiskers for IMP accuracy.

In order to explore the generality of the architecture for
different monitors, we implemented four different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
Note that, although our dataflow engine performs similar operations to the DIFT
monitor, the dataflow engine operates on smaller, 2-bit flags.

\ED{We need to point out that these monitors cannot be implemented with a simple
1-bit dataflow engine (DIFT) that we use for filtering. 
Point: address potential reviewer questions on why don't we just use a 
programmable dataflow engine for monitoring.}

We tested our system using all C benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. An initial slack of 2
million cycles, which is less than 1\% of
the total execution time, is given for a 10\% target overhead. This initial
slack is scaled proportionally for other overhead targets. Unless otherwise
specified, an unrestricted, slack-based dropping policy is used.

@@

% Full monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
    \vspace{-0.2in}
    \caption{Full monitoring overheads.}
    \label{fig:evaluation.full_mon}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.full_mon} shows the execution times of
performing full monitoring normalized to the execution times of the benchmarks
without monitoring. In these results, no filtering or partial monitoring is
done. UMC shows normalized execution times from 5x to 16x with an average
of 10x. BC shows normalized execution times of 8-24x with an average of
16x while DIFT shows normalized execution times of 4x-14x with an average of
9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
% These monitors show high overheads because they require several
% instructions to process each monitored instruction from the main core. Even
% small increases in the number of instructions needed by the monitor lead to
% high overheads when multiplied across the main program's execution.

\ED{we need to better justify the high overhead. 
Key points: 1) the overhead is in line with LBA, which represent multi-core
monitoring w/ low HW costs.
2) If we are going to mention dynamic memory allocation. We need to say why
it's really necessary -- it's not our implementation problem. It's something
that you need to have for large meta-data. Also, I don't think UMC would need
dynamic allocation.}

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with filtering null metadata.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 1.7-7x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 65\%. For BC, normalized execution
times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
times for filtered BC is 1.1-8.3x. 
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
on average after filtering, an 88\% reduction in monitoring overheads, with a
maximum overhead of 31\% after filtering. This is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering. IMP does not have
null metadata and thus is not affected by null metadata filtering.
These overheads are similar to previous results for multi-core monitoring with
minimal hardware costs such as LBA \cite{lba-asid06} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for array bounds check.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Error versus varying overhead budget for instruction-mix profiling.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
instruction types for each benchmark and for varying overheads. We see that
with a 1.1x overhead a max error of 9\% is observed and an average of less than
4\%. This error decreases as overhead is increased and is 0\% for all
benchmarks once a 6x overhead target is used.

@@

% Flex UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies with probabilistic dropping. Here, we use a 10\% probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that since there are multiple dependent
monitoring operations that must be done in order for a final check to be
successful as discussed in Section~\ref{sec:policies.when}.
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.

>>>>

In order to explore the generality of the architecture for
different monitors, we implemented four different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations. Additionally, recent
work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
LockSet attempts to detect possible race conditions in multi-threaded programs
by associating information about which locks have access to shared memory
locations. If a shared memory location is accessed without the use of a lock
then a race condition may exist.
Note that, although our dataflow engine performs similar operations to the DIFT
monitor, the dataflow engine operates on smaller, 2-bit flags.

We tested our system using benchmarks from SPECint
CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
modification of {\tt malloc} to set array bounds information, we focus on the C
SPECint benchmarks. Although we do not
show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
for these benchmarks are similar to the other results shown. For each
benchmark, we simulated for 2 billion instructions. 
% An initial slack of 2
% million cycles, which is less than 1\% of
% the total execution time, is given for a 1.1x target overhead. This initial
% slack is scaled proportionally for other overhead targets. 
Since LS race detection requires multi-threaded programs, we tested LS using
the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
200 million instructions.
Unless otherwise specified, an unrestricted, slack-based dropping policy is
used.

@@

% % Full monitoring overheads
% \begin{figure}
%   \begin{center}
%     \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
%     \vspace{-0.2in}
%     \caption{Full monitoring overheads.}
%     \label{fig:evaluation.full_mon}
%     \vspace{-0.1in}
%   \end{center}
% \end{figure}

% Figure~\ref{fig:evaluation.full_mon} shows the execution times of
% performing full monitoring normalized to the execution times of the benchmarks
% without monitoring. In these results, no filtering or partial monitoring is
% done. 
Our initial results for performing full core-based monitoring show severalfold slowdowns.
UMC shows normalized execution times from 5x to 16x with an average
of 10x. BC shows normalized execution times of 8-24x with an average of
16x while DIFT shows normalized execution times of 4x-14x with an average of
9x. IMP shows normalized execution times from 2-6x with an average of 4x and LS
shows slowdowns of 1.8-3.3x with an average of 2.4x.

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with filtering null metadata.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
normalized execution times with filtering enabled. We see significant
reductions in overhead for all three monitoring schemes. UMC sees normalized
execution times of 1.7-7x with an average of 4x with null metadata filtering.
This is an average reduction in overheads of 65\%. For BC, normalized execution
times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
times for filtered BC is 1.1-8.3x. 
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
on average after filtering, an 88\% reduction in monitoring overheads, with a
maximum overhead of 31\% after filtering. This is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. For most of these
benchmarks, this propagates to relatively few instructions. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would expect to see less filtering. IMP does not have
null metadata and thus is not affected by null metadata filtering.
LS (not shown) had 1.03-1.72x after null metadata filtering and average
overhead observed was 1.33x for the SPLASH-2 benchmarks.
These overheads are similar to previous results for multi-core monitoring with
minimal hardware costs such as LBA \cite{lba-asid06} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for BC.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for UMC.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% LS sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_ls_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for LS.}
    \label{fig:evaluation.ls_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.
Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
71\% coverage with a 1.2x overhead. 
\DAN{Try to explain water\_nsquared results not reaching 100\%?}

The instruction-mix profiling monitor does not perform check operations. Thus,
the concept of coverage is not applicable here. Instead, for each instruction
type profiled, we calculate its count as a percentage of the total instructions
monitored. We compare these percentages against the case when full monitoring
is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the min, max, and average error across
instruction types for each benchmark and for varying overheads. We see that
with a 1.1x overhead a max error of 9\% is observed across benchmarks and the
average error across instruction types and benchmarks is less than 2\%.

@@

% Flex UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_exec_time.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \vspace{-0.1in}
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

@@

\begin{figure}
  \begin{center}
  \vspace{-0.2in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_exec_time.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \vspace{-0.2in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \vspace{-0.1in}
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.2in 0 0]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One possible application of partial monitoring with low overheads is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overheads across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check running for
unrestricted and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped. Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored. Each run
was simulated for 200 million instructions. 
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that since there are multiple dependent
monitoring operations that must be done in order for a final check to be
successful as discussed in Section~\ref{sec:policies.when}.
Instead, source-only
dropping is much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 5\% increase in coverage with only seven runs.
Over hundreds or thousands of runs, source-only dropping should be able to
reach 100\% code coverage.
